---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(openxlsx)
library(xml2)
library(stats)
library(utils)
library(conflicted)
library(here)
library(car)
library(rstatix)
library(hpar)
library(gridExtra)
library(HPAanalyze)
```

# Quantity

```{r}
# Post-hoc analysis: Dunn's test
skinECM %>%
  dunn_test(Quantity~Age2) # sig difference between adult and the remaining two groups, but no sig difference between young and elderly
```

```{r}
# Wilcoxon test to identify genes with significantly different expression across different age groups
skinECM %>% semi_join(., count(., Gene_name, Age2) %>% dplyr::filter(n > 2)) %>% 
  semi_join(., count(., Gene_name, Age2) %>% count(Gene_name) %>% dplyr::filter(n > 1)) %>%
  group_by(Gene_name) %>%
  wilcox_test(Quantity~Age2) %>%
  dplyr::filter(p.adj < 0.05) %>%
  arrange(p)
```

```{r}
# Correlation test between intensity and age for every gene
skinECM %>% 
  semi_join(., count(., Gene_name, Age) %>% dplyr::filter(n > 2)) %>% 
  group_by(Gene_name) %>%
  cor_test(Quantity, Age) %>%
  arrange(desc(abs(cor))) %>%
  dplyr::filter(p < 0.05) 

# Plot of genes with significant correlation
skinECM %>%
  dplyr::filter(Gene_name %in% c("PDGFB", "PPA2", "IDH1", "UGDH", "DECR1", "ENG", "APOE", "APOA4")) %>%
  ggplot() + aes(x = Age, y = Quantity) + geom_point() + facet_wrap(~Gene_name, scales = "free_x") + geom_smooth(method = "lm")
```

```{r}
# Boxplot of expression of sig cor genes across age groups
skinECM %>%
  dplyr::filter(Gene_name %in% c("PDGFB", "PPA2", "IDH1", "UGDH", "DECR1", "ENG", "APOE", "APOA4")) %>%
  ggplot() + aes(x = Age2, y = Quantity) + geom_boxplot() + facet_wrap(~Gene_name) + scale_x_discrete(limits = level_order)
```

```{r}
# Genes of interest: DECR1, PDGFB, IDH1, APOA4, ENG

level_order <- c('young', 'adult', 'elderly')

skinECM %>%
  dplyr::filter(Gene_name %in% c("DECR1", "PDGFB", "IDH1", "APOA4", "ENG")) %>%
  ggplot() + aes(x = Age2, y = Quantity) + geom_boxplot() + facet_wrap(~Gene_name) + scale_x_discrete(limits = level_order)
```

# Genes of interest

```{r}
Histology <- subsetData$normal_tissue
Pathology <- subsetData$pathology
cellLocal <- subsetData$subcellular_location

Histology %>%
  dplyr::filter(gene %in% c("MMP9", "DECR1", "PDGFB", "IDH1", "APOA4", "ENG"))

Pathology %>%
  dplyr::filter(gene %in% c("MMP9", "DECR1", "PDGFB", "IDH1", "APOA4", "ENG"))

cellLocal %>%
  dplyr::filter(gene %in% c("MMP9", "DECR1", "PDGFB", "IDH1", "APOA4", "ENG"))
```

*ENG gene* -> encodes endoglin protein (https://www.proteinatlas.org/ENSG00000106991-ENG/tissue/skin)
- the only gene that is sig DE and sig corr for both parameters and ageing
```{r}
skinHistology %>%
  dplyr::filter(gene == "ENG")

hpaVis(data = subsetData,
       targetGene = 'ENSG00000106991',
       targetTissue = c('skin', 'skin 1', 'skin 2'),
       targetCellType = "cells in basal layer",
       visType = c("Tissue", "Subcell"))
```

*MMP9 gene* -> encodes matrix metallopeptidase 9 (https://www.proteinatlas.org/ENSG00000100985-MMP9/tissue/skin)
- sig DE and corr for intensity parameter and ageing -> low exp in Langerhans, melanocytes and keratinocytes
```{r}
skinHistology %>%
  dplyr::filter(gene == "MMP9")

hpaVis(data = subsetData,
       targetGene = 'ENSG00000100985',
       targetTissue = c('skin', 'skin 1', 'skin 2'),
       targetCellType = "cells in basal layer",
       visType = c("Tissue", "Subcell"))
```

*IDH1 gene* (https://www.proteinatlas.org/ENSG00000138413-IDH1/tissue/skin)
- sig DE and corr for quantity parameter -> high exp in sebaceous glands, low in melanocytes, keratinocytes and epidermal cells
```{r}
skinHistology %>%
  dplyr::filter(gene == "IDH1")

hpaVis(data = subsetData,
       targetGene = 'ENSG00000138413',
       targetTissue = c('skin', 'skin 1', 'skin 2'),
       targetCellType = "cells in basal layer",
       visType = c("Tissue", "Subcell"))
```

*DECR1 gene* (https://www.proteinatlas.org/ENSG00000104325-DECR1/tissue/skin)
```{r}
skinHistology %>%
  dplyr::filter(gene == "DECR1")

hpaVis(data = subsetData,
       targetGene = 'ENSG00000104325',
       targetTissue = c('skin', 'skin 1', 'skin 2'),
       targetCellType = "cells in basal layer",
       visType = c("Tissue", "Subcell"))
```

*PDGFB gene* (https://www.proteinatlas.org/ENSG00000100311-PDGFB/tissue/skin)
- ubiquitous cytoplasmic expression
```{r}
skinHistology %>%
  dplyr::filter(gene == "PDGFB")

hpaVis(data = subsetData,
       targetGene = 'ENSG00000100311',
       targetTissue = c('skin', 'skin 1', 'skin 2'),
       targetCellType = "cells in basal layer",
       visType = c("Tissue", "Subcell"))
```

*APOA4 gene* (https://www.proteinatlas.org/ENSG00000110244-APOA4)
```{r}
skinHistology %>%
  dplyr::filter(gene == "APOA4")

hpaVis(data = subsetData,
       targetGene = 'ENSG00000110244',
       targetTissue = c('skin', 'skin 1', 'skin 2'),
       targetCellType = "cells in basal layer",
       visType = c("Tissue", "Subcell"))
```

