---
title: "Visualization of the raw data"
output: html_notebook
---

```{r}
if (!require(devtools)) install.packages("devtools")
devtools::install_github("gaospecial/ggVennDiagram")

install.packages("VennDiagram")
```

```{r}
library(tidyverse)
library(openxlsx)
library(xml2)
library(utils)
library(conflicted)
library(here)
library(car)
library(rstatix)
library(ggVennDiagram)
library(VennDiagram)
library(grid)
library(futile.logger)
```

# Structure of the raw data
How many subjects (sexes), antibodies, genes and tissues are in the entire dataset?

## Subjects
```{r}
# Subjects divided by biological sex
level_order <- c('≤ 40', '> 60')

healthyECM %>%
  group_by(Age2) %>%
  count(Sex) %>%
  ggplot() + aes(x = Age2, y = n, colour = Sex, fill = Sex, label = n) +
  geom_col(position = position_dodge()) + 
  ylab("Frequency of subjects") + xlab("Age") +
  scale_x_discrete(limits = level_order) + theme_classic() + theme(
  axis.text = element_text(size = 12, colour = "black"),
  axis.title = element_text(size = 13),
  title = element_text(size = 14),
  legend.position = "top",
  plot.title = element_text(hjust = 0.5)
  ) +
  geom_text(position = position_dodge(width = 1), vjust = -0.5, size = 3) +
  ggtitle("Frequency of subjects divided by biological sex")

ggsave(here("output/rawData_sex.tiff"))

healthyECM %>%
  group_by(Age2) %>%
  get_summary_stats()

# Age distribution
ECM %>%
  ggplot() + aes(x = Age) + geom_density(colour = "#00BFC4") + theme_classic() + theme(
  plot.title = element_text(hjust = 0.5),
  axis.text = element_text(size = 12, colour = "black"),
  axis.title = element_text(size = 13),
  title = element_text(size = 15)) + ylab("Kernel density estimate of subject counts") + ggtitle("Age distribution in the raw data")
ggsave(here("output/rawData_ageDist.tiff"), height = 7, width = 9)

ECM %>%
  ggplot() + aes(x = Age) + geom_histogram(fill = "#00BFC4") + theme_classic() + theme(
  plot.title = element_text(hjust = 0.5),
  axis.text = element_text(size = 12, colour = "black"),
  axis.title = element_text(size = 13),
  title = element_text(size = 15)) + ggtitle("Age distribution in the raw data") + ylab("Count of subjects")
ggsave(here("output/rawData_ageDist_histogram.tiff"), height = 7, width = 9)

healthyECM %>%
  ggplot() + aes(x = Age) + geom_density(colour = "#00BFC4") + theme_classic() + theme(
  plot.title = element_text(hjust = 0.5),
  axis.text = element_text(size = 12, colour = "black"),
  axis.title = element_text(size = 13),
  title = element_text(size = 15)) + ylab("Kernel density estimate of subject counts") + ggtitle("Age distribution in the filtered data") + geom_vline(xintercept = 40, linetype = "dashed", colour = "gray", size = 0.5) + geom_vline(xintercept = 60, linetype = "dashed", colour = "gray", size = 0.5)
ggsave(here("output/filteredData_ageDist.tiff"), height = 7, width = 9)

healthyECM %>%
  ggplot() + aes(x = Age) + geom_histogram(fill = "#00BFC4") + theme_classic() + theme(
  plot.title = element_text(hjust = 0.5),
  axis.text = element_text(size = 12, colour = "black"),
  axis.title = element_text(size = 13),
  title = element_text(size = 15)) + ggtitle("Age distribution in the filtered data") + geom_vline(xintercept = 40, linetype = "dashed", colour = "gray", size = 0.5) + geom_vline(xintercept = 60, linetype = "dashed", colour = "gray", size = 0.5) + ylab("Count of subjects")
ggsave(here("output/filteredData_ageDist_histogram.tiff"), height = 7, width = 9)

```

## Antibodies
```{r}
# Antibodies
healthyECM %>%
  group_by(Age2) %>%
  count(Antibody_name)

x <- healthyECM %>%
  select(Age2, Antibody_name) %>%
  dplyr::filter(Age2 == "≤ 40") %>%
  pull()

x1 <- healthyECM %>%
  select(Age2, Antibody_name) %>%
  dplyr::filter(Age2 == "> 60") %>%
  pull()

venn(y, snames = c("≤ 40", "> 60"), zcolor = "#00BFC4", col = "#00BFC4", ilcs = 1.25, sncs = 1.25, borders = FALSE, box = FALSE, ggplot = TRUE)  

ggsave(here("output/antibodyVenn.tiff"))
```

## Genes
```{r}
x2 <- healthyECM %>%
  select(Age2, Gene_name) %>%
  dplyr::filter(Age2 == "≤ 40") %>%
  pull()

x3 <-  healthyECM %>%
  select(Age2, Gene_name) %>%
  dplyr::filter(Age2 == "> 60") %>%
  pull()

y1 <- list(x2, x3)

venn(y1, snames = c("≤ 40", "> 60"), zcolor = "#00BFC4", col = "#00BFC4", ilcs = 1.25, sncs = 1.25, borders = FALSE, box = FALSE, ggplot = TRUE)  

ggsave(here("output/genesVenn.tiff"))
```

## Tissues
```{r}
healthyECM %>%
  count(Tissue_type_1)

x4 <- healthyECM %>%
  select(Age2, Tissue_type_1) %>%
  dplyr::filter(Age2 == "≤ 40") %>%
  pull()

x5 <-  healthyECM %>%
  select(Age2, Tissue_type_1) %>%
  dplyr::filter(Age2 == "> 60") %>%
  pull()

y2 <- list(x4, x5)

venn(y2, snames = c("≤ 40", "> 60"), zcolor = "#00BFC4", col = "#00BFC4", ilcs = 1.25, sncs = 1.25, borders = FALSE, box = FALSE, ggplot = TRUE)  
# esophagus, placenta and tonsil not in both age groupings

ggsave(here("output/tissuesVenn.tiff"))
```

# Age distribution of the raw dataset and the two age-groups in this study
How many images are there in each age group? How many genes contain antibodies in both age groups?
```{r}
# How many images are there in each group?
healthyECM %>%
  dplyr::filter(Age2 == "≤ 40") %>%
  count(Image_file) # 44,552

healthyECM %>%
  dplyr::filter(Age2 == "> 60") %>%
  count(Image_file) # 45,447

# How many genes contain antibodies in both age groups?
healthyECM %>%
  group_by(Gene_name) %>%
  count(., Antibody_name, Age2) %>% 
  count(Antibody_name) %>% dplyr::filter(n > 1) %>%
  distinct() # all genes contain at least 1 antibody in both age groups
```

# Inferential stats visualization
Figures explaining the statistical tests I did.
```{r}

```

# DE
List all the genes that show significant difference between the two age groups in terms of intensity and quantity.
```{r}
wilcoxResults # sig DE genes (p < 0.05) found in 26 tissue types

# Sig DE genes in order of the most to least frequently occuring ones
wilcoxResults %>%
  ungroup() %>%
  count(Gene_name) %>%
  arrange(desc(n))
```

# Age-related genes
Summarise how many genes are found to be age-related in 1, 2, 3 and >3 tissues. About the HPA images, I think we need the images from one to two strong examples of known aging markers.
```{r}
# Intensity
iTissueCount <- wilcoxIntensity %>%
  ungroup() %>%
  count(Gene_name)

wilcoxIntensity <- full_join(wilcoxIntensity, iTissueCount)

wilcoxIntensity <- wilcoxIntensity %>%
  mutate(nTissues = case_when(
    n == 1 ~ "1",
    n == 2 ~ "2",
    n == 3 ~ "3",
    n > 3 ~ "> 3"
  ))

# Quantity
qTissueCount <- wilcoxQuantity %>%
  ungroup() %>%
  count(Gene_name)

wilcoxQuantity <- full_join(wilcoxQuantity, qTissueCount)

wilcoxQuantity <- wilcoxQuantity %>%
  mutate(nTissues = case_when(
    n == 1 ~ "1",
    n == 2 ~ "2",
    n == 3 ~ "3",
    n > 3 ~ "> 3"
  ))

wilcoxResults <- full_join(wilcoxIntensity, wilcoxQuantity)
```

```{r}
# Plot
level_order1 <- c('> 3', '3', '2', '1')

wilcoxResults %>%
  dplyr::filter(.y. == "Intensity") %>%
  ggplot() + aes(y = nTissues) + geom_bar(position = position_dodge(), orientation = "y", fill = "#00BFC4") + ggtitle("Age-related genes in one or more tissue types") + scale_y_discrete(limits = level_order1) + xlab("Number  of genes") + ylab("Number of tissue types") + labs(subtitle = "Significant difference in Intensity") + scale_x_continuous(position = "bottom") + theme_classic() + theme(
  axis.text = element_text(size = 12, colour = "black"),
  axis.title = element_text(size = 13),
  title = element_text(size = 15),
  plot.title = element_text(hjust = 0.5),
  plot.subtitle = element_text(hjust = 0.5)) 
  #scale_fill_manual(values = c("#00BCD4", "#FFC107"), na.translate = F)

ggsave(here("output/ageRelated_barchart_intensity.tiff"))

wilcoxResults %>%
  dplyr::filter(.y. == "Quantity") %>%
  ggplot() + aes(y = nTissues) + geom_bar(position = position_dodge(), orientation = "y", fill = "#00BFC4") + ggtitle("Age-related genes in one or more tissue types") + scale_y_discrete(limits = level_order1) + xlab("Number  of genes") + ylab("Number of tissue types") + labs(subtitle = "Significant difference in Quantity") + scale_x_continuous(position = "bottom") + theme_classic() + theme(
  axis.text = element_text(size = 12, colour = "black"),
  axis.title = element_text(size = 13),
  title = element_text(size = 15),
  plot.title = element_text(hjust = 0.5),
  plot.subtitle = element_text(hjust = 0.5)) 
  #scale_fill_manual(values = c("#00BCD4", "#FFC107"), na.translate = F)

ggsave(here("output/ageRelated_barchart_quantity.tiff"))
```



