---
title: "Automated loops"
output: html_notebook
---

```{r}
library(tidyverse)
library(openxlsx)
library(xml2)
library(utils)
library(conflicted)
library(here)
library(car)
library(rstatix)
library(purrr)
```

# Age classification in the entire dataset
```{r}
healthyECM <- ECM %>%
  dplyr::filter(Tissue_type_2 == "Normal tissue, NOS (M-00100)")

# Sample classification by age 
healthyECM$Age2 <- healthyECM$Age

healthyECM <- healthyECM %>%
  mutate(Age2 = case_when(
    Age <= 40 ~ "≤ 40",
    Age > 60 ~ "> 60"
  ))

healthyECM <- healthyECM %>%
  drop_na(Age2)

healthyECM %>%
  count(Age2)

## Sample classification by age: take a % from each age grouping and make it into a table
ageData <- healthyECM %>%
  group_by(Tissue_type_1) %>%
  count(Age2) %>%
  mutate(Percentage = n / sum(n) * 100) %>%
  select(Tissue_type_1, Age2, Percentage) %>%
  pivot_wider(names_from = Age2, values_from = Percentage) # - esophagus, placenta, tonsil 

healthyECM <- healthyECM %>%
  dplyr::filter(!Tissue_type_1 == "placenta") %>%
  dplyr::filter(!Tissue_type_1 == "esophagus") %>%
  dplyr::filter(!Tissue_type_1 == "tonsil")

healthyECM %>%
  count(Tissue_type_1)
```

```{r}
# This plot has to go into the dissertation/paper -> explains why the age was classified the way it was (thirds)
healthyECM %>%
  count(Age)

healthyECM %>%
  ggplot() + aes(x = Age) + geom_density()

healthyECM %>%
  get_summary_stats(Age)
```

```{r}
# Tesf of normality: density plots
## Intensity distribution
healthyECM %>%
  select(Age2, Intensity) %>%
  ggplot() + aes(x = Intensity, colour = Age2) + geom_density() # not a normal distribution -> positively skewed

## Quantity distribution
healthyECM %>%
  select(Age2, Quantity) %>%
  ggplot() + aes(x = Quantity, colour = Age2) + geom_density() # not a normal distribution -> VERY positively skewed
```

# Automate the following functions for all datasets
## Non-parametric testing: Kruskal-Wallis -> is there a significant difference between the samples?
```{r}
# Intensity
doKruskalTest <- function(x) {
  x %>%
    kruskal_test(Intensity~Age2) %>%
  print(kruskalTest)
}

a <- healthyECM %>%
  group_by(Tissue_type_1) %>%
  nest()

a$data[[1]]

a <- healthyECM %>%
  group_by(Tissue_type_1) %>%
  nest() %>%
  mutate(Kruskal = map(data, doKruskalTest))

a$Kruskal[[1]]

# Quantity
doKruskalTest1 <- function(x) {
  x %>%
    kruskal_test(Quantity~Age2) %>%
  print(kruskalTest1)
}

b <- healthyECM %>%
  group_by(Tissue_type_1) %>%
  nest()

b$data[[1]]

b <- healthyECM %>%
  group_by(Tissue_type_1) %>%
  nest() %>%
  mutate(Kruskal = map(data, doKruskalTest1))

b$Kruskal[[1]]

# Tidy the data frames
a <- a %>%
  unnest()

b <- b %>%
  unnest()

kruskalResults <- full_join(a, b) %>%
  select(Tissue_type_1, .y., n, statistic, df, p, method) %>%
  unique() %>%
  dplyr::filter(p < 0.05) %>%
  group_by(Tissue_type_1) %>%
  arrange(.by_group = TRUE)
kruskalResults

kruskalResults %>%
  count(Tissue_type_1) # 37 tissues have sig difference

healthyECM %>%
  count(., Gene_name, Age2) %>% 
  count(., Gene_name, Age2) %>% count(Gene_name) %>% dplyr::filter(n > 1) # all genes are present in both age groups in all tissues
```

## Post-hoc analysis: Dunn's test -> which samples significantly statistically differ?
```{r}
# Intensity
doDunnsTest <- function(x) {
    x %>%
    dunn_test(Intensity~Age2) %>%
  print(dunnsTest)
  }

c <- healthyECM %>%
  group_by(Tissue_type_1) %>%
  nest()

c$data[[1]]

c <- healthyECM %>%
  group_by(Tissue_type_1) %>%
  nest() %>%
  mutate(Dunn = map(data, doDunnsTest))

c$Dunn[[1]]

# Quantity
doDunnsTest1 <- function(x) {
    x %>%
    dunn_test(Quantity~Age2) %>%
  print(dunnsTest1)
  }

d <- healthyECM %>%
  group_by(Tissue_type_1) %>%
  nest()

d$data[[1]]

d <- healthyECM %>%
  group_by(Tissue_type_1) %>%
  nest() %>%
  mutate(Dunn = map(data, doDunnsTest1))

d$Dunn[[1]]

# Tidy the data frames
c <- c %>%
  unnest(Tissue_type_1, Dunn)

d <- d %>%
  unnest(Tissue_type_1, Dunn)

dunnResults <- full_join(c, d) %>%
  select(Tissue_type_1, .y., group1, group2, n1, n2, statistic, p, p.adj, p.adj.signif) %>%
  dplyr::filter(p.adj < 0.05) %>%
  group_by(Tissue_type_1) %>%
  arrange(.by_group = TRUE)
dunnResults
```

## Wilcoxon test: which genes are significantly DE across different age groupings?
```{r}
# Intensity
doWilcoxTest <- function(x) {
  x %>%
    semi_join(., count(., Gene_name, Age2)) %>% 
    semi_join(., count(., Gene_name, Age2) %>% count(Gene_name) %>% dplyr::filter(n > 1)) %>%
    group_by(Gene_name) %>%
    wilcox_test(Intensity~Age2) %>%
  print(WilcoxTest)
  }

e <- healthyECM %>%
  group_by(Tissue_type_1) %>%
  nest()

e$data[[18]]

e <- healthyECM %>%
  group_by(Tissue_type_1) %>%
  nest() %>%
  mutate(Wilcox = map(data, doWilcoxTest))

e$Wilcox[[1]]

# Quantity
doWilcoxTest1 <- function(x) {
  x %>%
    semi_join(., count(., Gene_name, Age2)) %>% 
    semi_join(., count(., Gene_name, Age2) %>% count(Gene_name) %>% dplyr::filter(n > 1)) %>%
    group_by(Gene_name) %>%
    wilcox_test(Quantity~Age2) %>%
  print(WilcoxTest1)
  }

f <- healthyECM %>%
  group_by(Tissue_type_1) %>%
  nest()

f$data[[1]]

f <- healthyECM %>%
  group_by(Tissue_type_1) %>%
  nest() %>%
  mutate(Wilcox = map(data, doWilcoxTest1))

f$Wilcox[[1]]

# Tidy the data frames
e <- e %>%
  unnest(Tissue_type_1, Wilcox)

f <- f %>% 
  unnest(Tissue_type_1, Wilcox)

wilcoxResults <- full_join(e, f) %>%
  select(Tissue_type_1, Gene_name, .y., group1, group2, n1, n2, statistic, p) %>%
  dplyr::filter(p < 0.05)
wilcoxResults
```

## Correlation test: is there a correlation between aging and the significantly DE genes?
```{r}
corData <- inner_join(wilcoxResults, healthyECM)
corData$Unknown_code_2 <- NULL

# Intensity
doCorrTest <- function(x) {
  x %>%
    semi_join(., count(., Gene_name, Age)) %>% 
    group_by(Gene_name) %>%
    cor_test(Intensity, Age, method = "spearman") %>%
    print(CorrTest)
}

g <- corData %>%
  group_by(Tissue_type_1) %>%
  dplyr::filter(Tissue_type_1 != "bone+marrow") %>%
  nest()

g$data[[3]]

g <- corData %>%
  group_by(Tissue_type_1) %>%
  nest() %>%
  mutate(Correlation = map(data, doCorrTest))

#g$Correlation[[1]]

# Quantity
doCorrTest1 <- function(x) {
  x %>%
    semi_join(., count(., Gene_name, Age)) %>% 
    group_by(Gene_name) %>%
    cor_test(Quantity, Age, method = "spearman") %>%
    print(CorrTest1)
}

h <- healthyECM %>%
  group_by(Tissue_type_1) %>%
  nest()

h$data[[1]]

h <- healthyECM %>%
  group_by(Tissue_type_1) %>%
  nest() %>%
  mutate(Correlation = map(data, doCorrTest1))

h$Correlation[[1]]

# Tidy the data frames
g <- g %>%
  unnest(Tissue_type_1, Correlation)

h <- h %>%
  unnest(Tissue_type_1, Correlation)

corrResults <- full_join(g, h) %>%
  select(Tissue_type_1, Gene_name, var1, var2, cor, statistic, p, method) %>%
  drop_na() %>%
  dplyr::filter(p < 0.05)
corrResults
```








